\documentclass[17pt]{beamer}
\usetheme{Warsaw}

\title{XMPP testing with Escalus}
\author{Krzysztof Goj}

\usepackage{minted}
\newminted{erlang}{gobble=2,fontsize=\small,baselinestretch=1.0}

\begin{document}


\begin{frame}
\titlepage
\end{frame}

\begin{frame}{What is Escalus?}
\begin{center}
\textbf{Escalus is \alert<1>{a library} \alert<2>{for acceptance testing} \alert<3>{XMPP servers}.}

\vspace{1em}

\begin{footnotesize}
\only<1>{some code that makes doing certain things \textbf{easier}}
\only<2>{checking if a thing \textbf{does what it's meant to}}
\only<3>{XMPP (Jabber) - eXtensible Messaging and Presence Protocol}
\end{footnotesize}
\end{center}
\end{frame}

\begin{frame}{Background}
\begin{small}
\begin{itemize}
\item XMPP is an extensible protocol
\item Our job is often to extend or adjust ejabberd\\ to fit customer's demand.
\item Ejabberd is a generic software - optimize by throwing away stuff you don't need.
\item How to make sure we did the right thing?
\item How to know if we break something as we go?
\end{itemize}
\end{small}
\end{frame}

\begin{frame}{Before Escalus}
Boring stuff is boring:
\begin{itemize}
\item Use Client's (say, Psi's) XMPP console.
\item Works good for simple scenarios
\item PITA with more complicated scenarios:
    \begin{itemize}
    \item several users,
    \item several resources per user,
    \item setting up \& tearing down the state.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{after Escalus}
Automated testing makes life easier:
\begin{itemize}
\item Continuous Integration
\item TDD
\item Set some tracing and replay the test.
\item Get Wireshark dump -- write a testcase.
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Escalus test case}
\begin{minted}[fontsize=\footnotesize]{erlang}
% Alice sends a chat message to Bob's bare JID
% Bob gets the message on both resources
% Bob replies to one of Alice's resources
% Alice receives the reply only on that resource
\end{minted}
\end{frame}

\begin{frame}[fragile]{Escalus test case}
\begin{minted}[fontsize=\tiny]{erlang}
message_routing(Config) ->
    escalus:story(Config, [2, 2],
        fun(Alice1, Alice2, Bob1, Bob2) ->
            % Alice sends a chat message to Bob's bare JID
            escalus:send(Alice1, escalus_stanza:chat_to(bob, <<"Coffee?">>),

            % Bob gets the message on both resources
            escalus:assert(is_chat_message, [<<"Coffee?">>],
                           escalus:wait_for_stanza(Bob1)),
            escalus:assert(is_chat_message, [<<"Coffee?">>],
                           escalus:wait_for_stanza(Bob2)),

            % Bob replies to one of Alice's resources
            escalus:send(Bob1, escalus_stanza:chat_to(Alice1, <<"Sure!">>),

            % Alice receives the reply only on that resource
            escalus:assert(is_chat_message, [<<"Sure!">>],
                           escalus:wait_for_stanza(Alice1)),
            escalus_assert:has_no_stanzas(Alice2)
        end).
\end{minted}
\end{frame}

\begin{frame}[fragile]{Think of it as a DSL}
\begin{minted}[fontsize=\tiny]{erlang}
message_routing(Config) ->
    escalus:story(Config, [2, 2], fun(Alice1, Alice2, Bob1, Bob2) ->
        % Alice sends a chat message to Bob's bare JID
        escalus:send(Alice1, escalus_stanza:chat_to(bob, <<"Coffee?">>),

        % Bob gets the message on both resources
        escalus:assert(is_chat_message, [<<"Coffee?">>],
                       escalus:wait_for_stanza(Bob1)),
        escalus:assert(is_chat_message, [<<"Coffee?">>],
                       escalus:wait_for_stanza(Bob2)),

        % Bob replies to one of Alice's resources
        escalus:send(Bob1, escalus_stanza:chat_to(Alice1, <<"Sure!">>),

        % Alice receives the reply only on that resource
        escalus:assert(is_chat_message, [<<"Sure!">>],
                       escalus:wait_for_stanza(Alice1)),
        escalus_assert:has_no_stanzas(Alice2)
    end).
\end{minted}
\end{frame}

\begin{frame}[fragile]{Common Test config file}
\begin{minted}[fontsize=\footnotesize]{erlang}
{escalus_users, [
    {alice, [
        {username, <<"alice">>},
        {server, <<"localhost">>},
        {password, <<"makota">>}]},
    {bob, [
        {username, <<"bob">>},
        {server, <<"localhost">>},
        {password, <<"mapsa">>}]}]}.
\end{minted}
\end{frame}

\begin{frame}{test results}
\begin{itemize}
\item \href{https://ci.erlang-solutions.com/job/ejabberd-esl/19/CT_Report/}{bad one}
\item \href{https://ci.erlang-solutions.com/job/ejabberd-esl/20/CT_Report/}{good one}
\end{itemize}
\end{frame}


\begin{frame}{What is done for us}
\begin{itemize}
\item User registration \& de-registration.
\item User login \& logout.
\item XML parsing \& generation.
\end{itemize}
\end{frame}

\begin{frame}{What is made easier}
\begin{itemize}
\item Checking assertions.
\item Debugging.
\end{itemize}
\end{frame}

\begin{frame}{Lesson learned}
\textbf{The Good:}
\begin{itemize}
\item acceptance testing,
\item ease of writing readable tests,
\item stories!
\end{itemize}
\end{frame}

\begin{frame}{Lessons learned}
\textbf{The Bad:} race conditions
\end{frame}

\begin{frame}{Lessons learned}
\textbf{The Ugly:}
\begin{itemize}
\item exmpp,
\item XML,
\item Common Test (a little).
\end{itemize}
\end{frame}

\begin{frame}{Short story of internals rewrite}
exmpp $\rightarrow$ exml + lxmppc
\begin{itemize}
\item internal rewrite
\item type madness: atoms, strings, binaries
\item warn \& go
\end{itemize}
\end{frame}

\begin{frame}{Future}
\begin{itemize}
\item BOSH support
\item property-based testing
\end{itemize}
\end{frame}

\begin{frame}{Summary}
\begin{itemize}
\item Think what to test.
\item Make writing tests easier.
\item Use Escalus for XMPP-related stuff.
\end{itemize}
\end{frame}

\begin{frame}{That's it!}
\begin{center}
\vspace{3em}

\textbf{\large Thank you for your attention!}

\vspace{2em}

\small
\href{mailto:krzysztof.goj@erlang-solutions.com}{krzysztof.goj@erlang-solutions.com}
\url{https://github.com/goj/escalus}
\end{center}
\end{frame}
\end{document}
